Bayesian Hyper-Heuristics

### 2023-02-04 article/sections/06-bayesian-hyper-heuristics.tex
🚧 (article): Save point

The predictive model thus models the \textit{proportional} probability of the event (selection of) heuristic $\boldsymbol{H}$, given allocation to entity $\boldsymbol{E}$ and credit requirement $\boldsymbol{C}$, parameterised by sampled $\boldsymbol{\theta} \vert \boldsymbol{\alpha} \sim Dir(\boldsymbol{\alpha}; K)$, $\boldsymbol{\phi} \vert \boldsymbol{\beta} \sim Dir(\boldsymbol{\beta}; K)^{J}$ and $\boldsymbol{\psi} \vert \gamma_{1}, \gamma_{0}  \sim Beta(\gamma_{1}, \gamma_{0})$. In the aforementioned, $K$ is the heuristic pool size, $J$ is the entity pool size and the indices $1$ and $0$ refers to success and failure parameters respectively. The concentration parameters $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$, $\gamma_{1}$ and $\gamma_{0}$ denote the prior beliefs of the \acs{BHH}. Similar to before, the reader is referred to \cite{ref:schreuder:2022} for the mathematical derivation of the selection mechanism.

@@

@@

Another approach to optimise the values for $\hat{\theta}_{k}$, $\hat{\phi}_{j,k}$ and $\hat{\psi}_{k}$ is to optimise the parameters by their probability distributions' parameters. This process is referred to as \textit{Bayesian analysis}. Bayesian analysis makes use of the \textit{posterior} probability distribution. The concentration update operations yielded by \acs{MAP}, are then be given as

@@

The following hyper-parameters are implemented by the \acs{BHH}: the \textit{heuristic pool} configures the type of heuristics included in the heuristic pool as gradient-only, meta-heuristic-only or all heuristics, the \textit{population size} specifies the number of entities in the entitiy pool, the \textit{credit assignment strategy} specifies which credit assignment strategy to use, the \textit{reselection interval} determines the frequency of heuristic reselection, the \textit{replay window size} determines the maximum size of the performance log, the \textit{reanalysis interval} determines the frequency at which Bayesian analysis is applied, the \textit{burn in window size} determines the size of an initial window where experience is simply gathered without reanalysis, and finally, \textit{discounted rewards} and \textit{normalisation} flags toggle scaling modifiers on the credit assignment strategies, backwards in the performance log.

>>>>

The predictive model thus models the \textit{proportional} probability of the event (selection of) heuristic $\boldsymbol{H}$, given allocation to entity $\boldsymbol{E}$ and credit requirement $\boldsymbol{C}$, parameterised by sampled $\boldsymbol{\theta} \sim Dir(\boldsymbol{\alpha}; K)$, $\boldsymbol{\phi} \sim Dir(\boldsymbol{\beta}; K)^{J}$ and $\boldsymbol{\psi} \sim Beta(\gamma_{1}, \gamma_{0})$. In the aforementioned, $K$ is the heuristic pool size and $J$ is the entity pool size. The concentration parameters $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$, $\gamma_{1}$ and $\gamma_{0}$ denote the prior beliefs of the \acs{BHH}. Similar to before, the reader is referred to \cite{ref:schreuder:2022} for the mathematical derivation of the selection mechanism.

@@

where $N_{k}$ is a summary variable denoting the count of occurrences of heuristic $k$, $N_{j}$ is a summary variable denoting the count of occurrences of entity $j$, $N_{j,k}$ is a summary variable denoting the count of occurrences of heuristic $k$ for entity $j$, and $N_{1,k}$ is a summary variable denoting the count of occurrences of heuristic $k$ that meet performance requirements (credit assignment), and $N$ is a summary variable denoting the total count of events.

@@

Another approach to optimise the values for $\hat{\theta}_{k}$, $\hat{\phi}_{j,k}$ and $\hat{\psi}_{k}$ is to optimise the parameters by their probability distributions' parameters. This process is referred to as \textit{Bayesian analysis}. Bayesian analysis makes use of the \textit{posterior} probability distribution. The concentration update operations yielded by \acs{MAP}, are then be given as follows:

@@

The following hyper-parameters are implemented by the \acs{BHH}: the \textit{heuristic pool} configures the type of heuristics included in the heuristic pool as gradient-only, meta-heuristic-only or all heuristics, the \textit{population size} specifies the number of entities in the entity pool, the \textit{credit assignment strategy} specifies which credit assignment strategy to use, the \textit{reselection interval} determines the frequency of heuristic reselection, the \textit{replay window size} determines the maximum size of the performance log, the \textit{reanalysis interval} determines the frequency at which Bayesian analysis is applied, the \textit{burn in window size} determines the size of an initial window where experience is simply gathered without reanalysis, and finally, \textit{discounted rewards} and \textit{normalisation} flags toggle scaling modifiers on values assigned by the credit assignment strategies, backwards in the performance log.

### 2023-02-04 article/sections/07-methodology.tex
🚧 (article): Save point

The \acs{BHH} baseline is a name given to a specific configuration of the \acs{BHH} which has been found to provide a reasonable baseline performance. The baseline configuration is used as the cornerstone configuration from which all other heuristics and their configurations are evaluated. The \acs{BHH} baseline configuration is given in Table \ref{tab:methodology:bhh_baseline_configuration}. In Table \ref{tab:methodology:bhh_baseline_configuration}, the heuristic pool configuration, \textit{all}, refers to a configuration where the heuristic pool contains all the low-level heuristics including all gradient-based heuristics and \acp{MH}.

% Table generated by Excel2LaTeX from sheet 'BHH Baseline'
\begin{table}[htb]
	\centering
	\caption{The \acs{BHH} baseline configuration as it is used in the empirical study.}
	\label{tab:methodology:bhh_baseline_configuration}%
	\par\bigskip
	\resizebox{0.65\textwidth}{!}{
		\begin{tabular}{rrcc}
			\multicolumn{1}{c}{\textbf{hyper-heuristic}} & \multicolumn{1}{c}{\textbf{variant}} & \textbf{configuration} & \textbf{value} \\
			\midrule
			\multicolumn{1}{l}{bhh}                      & \multicolumn{1}{l}{baseline}         & heuristic pool         & all            \\
			                                             &                                      & population             & 5              \\
			                                             &                                      & credit                 & ibest          \\
			                                             &                                      & burn in                & 0              \\
			                                             &                                      & reselection            & 10             \\
			                                             &                                      & replay                 & 10             \\
			                                             &                                      & reanalysis             & 10             \\
			                                             &                                      & normalise              & false          \\
			                                             &                                      & discounted rewards     & false          \\
		\end{tabular}%
	}
\end{table}%

@@

@@

@@

@@

All models that are trained in the empirical process follow implementations of shallow \acp{FFNN} with only one hidden layer. The number of hidden units used were determined empirically. Weights are initialised by means of Glorot uniform sampling \cite{ref:glorot:2010}. The models and their configuration, as it is used for each dataset, are given in Table \ref{tab:methodology:models:configurations}.

@@

\acs{BinXE} is used for classification problems with two classes and \acs{SparseCatXE} is used for classification problems with more than two classes. For the classification problems, accuracy is also measured. For regression problems, the \acs{MSE} is used as a performance metric. After training has been completed, the \textit{average rank}, based on test loss, for all configurations, is calculated at each mini-batch step.

@@

Each experiment and configuration is trained for a maximum of 30 epochs and is repeated over 30 independent runs, for each of the applicable datasets. Statistical analysis is executed on the results from the testing datasets. An average rank is calculated across all 30 runs, for both experimental groups and configurations, at each step, for every epoch of training.

@@

\subsection{Implementation and Execution}\label{sec:methodology:implementation}

>>>>

The \acs{BHH} baseline is a name given to a specific configuration of the \acs{BHH} which has been found to provide a reasonable baseline performance. The baseline configuration is used as the cornerstone configuration from which all other heuristics and their configurations are evaluated. The \acs{BHH} baseline configuration is given in Table \ref{tab:methodology:bhh_baseline_configuration}. In Table \ref{tab:methodology:bhh_baseline_configuration}, the heuristic pool configuration, \textit{all}, refers to a configuration where the heuristic pool contains all the low-level heuristics, including all gradient-based heuristics and \acp{MH}.

% Table generated by Excel2LaTeX from sheet 'BHH Baseline'
\begin{table}[htb]
	\centering
	\caption{The \acs{BHH} baseline configuration as it is used in the empirical study.}
	\label{tab:methodology:bhh_baseline_configuration}%
	\par\bigskip
	\resizebox{0.70\textwidth}{!}{
		\begin{tabular}{rrcc}
			\multicolumn{1}{c}{\textbf{hyper-heuristic}} & \multicolumn{1}{c}{\textbf{variant}} & \textbf{configuration} & \textbf{value} \\
			\midrule
			\multicolumn{1}{l}{bhh}                      & \multicolumn{1}{l}{baseline}         & heuristic pool         & all            \\
			                                             &                                      & population             & 5              \\
			                                             &                                      & credit                 & ibest          \\
			                                             &                                      & burn in                & 0              \\
			                                             &                                      & reselection            & 10             \\
			                                             &                                      & replay                 & 10             \\
			                                             &                                      & reanalysis             & 10             \\
			                                             &                                      & normalise              & false          \\
			                                             &                                      & discounted rewards     & false          \\
		\end{tabular}%
	}
\end{table}%

@@

\begin{table}[htb]
	\centering
	\caption{Low-level heuristics and their hyper-parameter configurations.}
	\label{tab:methodology:heuristics}%
	\par\bigskip
	\resizebox{0.7\textwidth}{!}{

		\begin{tabular}{llll}
			\textbf{heuristic} & \textbf{configuration}    & \textbf{value} & \textbf{citation}         \\
			\midrule
			sgd                & learning rate             & 0.1 (0.01)     & \cite{ref:sutskever:2013} \\
			momentum           & learning rate             & 0.1 (0.01)     & \cite{ref:sutskever:2013} \\
			                   & momentum                  & 0.9            &                           \\
			nag                & learning rate             & 0.1 (0.01)     & \cite{ref:sutskever:2013} \\
			                   & momentum                  & 0.9            &                           \\
			adagrad            & learning rate             & 0.1 (0.01)     & \cite{ref:duchi:2011}     \\
			                   & epsilon                   & 1E-07          &                           \\
			rmsprop            & learning rate             & 0.1 (0.01)     & \cite{ref:hinton:2012}    \\
			                   & rho                       & 0.95           &                           \\
			                   & epsilon                   & 1E-07          &                           \\
			adadelta           & learning rate             & 1.0 (0.95)     & \cite{ref:zeiler:2012}    \\
			                   & rho                       & 0.95           &                           \\
			                   & epsilon                   & 1E-07          &                           \\
			adam               & learning rate             & 0.1 (0.01)     & \cite{ref:kingma:2014}    \\
			                   & beta1                     & 0.9            &                           \\
			                   & beta2                     & 0.999          &                           \\
			                   & epsilon                   & 1E-07          &                           \\
			pso                & population size           & 10             & \cite{ref:van:2010}       \\
			                   & learning rate             & 1.0 (0.9)      &                           \\
			                   & inertia weight (w)        & 0.729844       &                           \\
			                   & cognitive control (c1)    & 1.49618        &                           \\
			                   & social control (c2)       & 1.49618        &                           \\
			                   & velocity clip min         & -1.0           &                           \\
			                   & velocity clip max         & 1.0            &                           \\
			de                 & population size           & 10             & \cite{ref:mezura:2006}    \\
			                   & selection strategy        & best           &                           \\
			                   & xo strategy               & exp            &                           \\
			                   & recombination probability & 0.9 (0.1)      &                           \\
			                   & beta                      & 2.0 (0.1)      &                           \\
			ga                 & population size           & 10             & \cite{ref:lambora:2019}   \\
			                   & selection strategy        & rand           &                           \\
			                   & xo strategy               & bin            &                           \\
			                   & mutation rate             & 0.2 (0.05)     &                           \\
		\end{tabular}%
	}
\end{table}%

@@

\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.8\textwidth]{images/bhh_heuristic_proxies.pdf}
	\caption{Mapping of proxied heuristic state update operations as implemented by the \acs{BHH}}
	\label{fig:methodology:heuristics:proxies}%
\end{figure}

@@

\begin{table}[htb]
	\centering
	\caption{Classification datasets}
	\label{tab:methodology:datasets:classification}%
	\par\bigskip
	\resizebox{\textwidth}{!}{
		\begin{tabular}{ccccccccc}
			\textbf{dataset} & \textbf{output} & \textbf{types}             & \textbf{attributes} & \textbf{classes} & \textbf{instances} & \textbf{batch} & \textbf{steps} & \textbf{citation}         \\
			\midrule
			iris             & multivariate    & real                       & 4                   & 3                & 150                & 16             & 10             & \cite{ref:fisher:1936}    \\
			car              & multivariate    & categorical                & 6                   & 4                & 1728               & 128            & 14             & \cite{ref:bohanec:1988}   \\
			abalone          & multivariate    & categorical, integer, real & 8                   & 28               & 4177               & 256            & 17             & \cite{ref:waugh:1995}     \\
			wine quality     & multivariate    & real                       & 12                  & 11               & 4898               & 256            & 20             & \cite{ref:cortez:2009}    \\
			mushroom         & multivariate    & categorical                & 22                  & 2                & 8214               & 512            & 17             & \cite{ref:schlimmer:1987} \\
			bank             & multivariate    & real                       & 17                  & 2                & 45211              & 512            & 89             & \cite{ref:moro:2014}      \\
			diabetic         & multivariate    & integer                    & 55                  & 3                & 100000             & 1024           & 98             & \cite{ref:strack:2014}    \\
		\end{tabular}%
	}
\end{table}%

\begin{table}[htb]
	\centering
	\caption{Regression datasets}
	\label{tab:methodology:datasets:regression}%
	\par\bigskip
	\resizebox{\textwidth}{!}{
		\begin{tabular}{cccccccc}
			\textbf{dataset}    & \textbf{output}           & \textbf{types} & \textbf{attributes} & \textbf{instances} & \textbf{batch} & \textbf{steps} & \textbf{citation}        \\
			\midrule
			fish toxicity       & multivariate              & real           & 7                   & 908                & 64             & 15             & \cite{ref:cassotti:2015} \\
			housing             & univariate                & real           & 13                  & 506                & 32             & 16             & \cite{ref:harrison:1978} \\
			forest fires        & multivariate              & real           & 13                  & 517                & 32             & 17             & \cite{ref:cortez:2007}   \\
			student performance & multivariate              & integer        & 33                  & 649                & 32             & 21             & \cite{ref:cortez:2008}   \\
			parkinsons          & multivariate              & integer, real  & 26                  & 5875               & 256            & 23             & \cite{ref:tsanas:2009}   \\
			air quality         & multivariate, time series & real           & 15                  & 9358               & 256            & 37             & \cite{ref:de:2008}       \\
			bike                & univariate                & integer, real  & 16                  & 17389              & 256            & 68             & \cite{ref:fanaee:2014}   \\
		\end{tabular}%
	}
\end{table}%

@@

All models that are trained in the empirical process follow implementations of shallow \acp{FFNN} with only one hidden layer. The number of hidden units used were determined empirically. Weights are initialised by means of \textit{Glorot uniform sampling} \cite{ref:glorot:2010}. The models and their configuration, as it is used for each dataset, are given in Table \ref{tab:methodology:models:configurations}.

\begin{table}[htb]
	\centering
	\caption{Model configurations}
	\label{tab:methodology:models:configurations}%
	\par\bigskip
	\resizebox{\textwidth}{!}{
		\begin{tabular}{rcccccccc}
			\textbf{dataset}    & \textbf{inputs} & \textbf{hidden} & \textbf{output} & \textbf{biases} & \textbf{parameters} & \textbf{topology} & \textbf{l1 activation} & \textbf{l2 activation} \\
			\midrule
			fish toxicity       & 6               & 3               & 1               & yes             & 25                  & dense             & LReLU ($\alpha = 0.3$) & sigmoid                \\
			iris                & 4               & 5               & 3               & yes             & 43                  & dense             & LReLU ($\alpha = 0.3$) & softmax                \\
			air quality         & 12              & 8               & 1               & yes             & 113                 & dense             & LReLU ($\alpha = 0.3$) & sigmoid                \\
			housing             & 13              & 8               & 1               & yes             & 121                 & dense             & LReLU ($\alpha = 0.3$) & sigmoid                \\
			wine quality        & 13              & 10              & 7               & yes             & 217                 & dense             & LReLU ($\alpha = 0.3$) & softmax                \\
			parkinsons          & 21              & 10              & 1               & yes             & 231                 & dense             & LReLU ($\alpha = 0.3$) & sigmoid                \\
			car                 & 21              & 10              & 4               & yes             & 264                 & dense             & LReLU ($\alpha = 0.3$) & softmax                \\
			forest fires        & 43              & 16              & 1               & yes             & 721                 & dense             & LReLU ($\alpha = 0.3$) & sigmoid                \\
			abalone             & 10              & 36              & 28              & yes             & 1432                & dense             & LReLU ($\alpha = 0.3$) & softmax                \\
			bank                & 51              & 32              & 1               & yes             & 1697                & dense             & LReLU ($\alpha = 0.3$) & softmax                \\
			bike                & 61              & 32              & 1               & yes             & 2017                & dense             & LReLU ($\alpha = 0.3$) & sigmoid                \\
			student performance & 99              & 32              & 1               & yes             & 3233                & dense             & LReLU ($\alpha = 0.3$) & sigmoid                \\
			adult               & 108             & 64              & 1               & yes             & 7041                & dense             & LReLU ($\alpha = 0.3$) & softmax                \\
			mushroom            & 117             & 64              & 1               & yes             & 7617                & dense             & LReLU ($\alpha = 0.3$) & softmax                \\
			diabetic            & 2369            & 32              & 3               & yes             & 75939               & dense             & LReLU ($\alpha = 0.3$) & softmax                \\
		\end{tabular}%
	}
\end{table}%

@@

\Acf{BinXE} is used for classification problems with two classes and \acf{SparseCatXE} is used for classification problems with more than two classes. For the classification problems, accuracy is also measured. For regression problems, the \acf{MSE} is used as a performance metric. After training has been completed, the \textit{average rank}, based on test loss, for all configurations, is calculated at each mini-batch step.

@@

Each experiment and configuration is trained for a maximum of 30 epochs and is repeated over 30 independent runs, for each of the applicable datasets. No early-stopping mechanism is used. Statistical analysis is executed on the results from the test datasets. An average rank is calculated across all 30 runs, for both experimental groups and configurations, at each step, for every epoch of training.

@@

\subsection{Implementation}\label{sec:methodology:implementation}
